A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-09 ISMB/ResponseToReviewers-Incarnation-ISMB-2013.tex

\Comment{What is the time consumption/complexity of the GC-weight parameter $x_{gc}$ identification within Sec. 2.2.3? And are these values length-/structure-independent or do they have to be computed for each structure individually?}
\Answer{These values are structure/\GCContent-dependent, and must be recomputed for any new input. We made this point clearer in the description of the adaptive sampling procedure, and the influence on the complexity of the number $M$ of iterations. Typically, 4 or 5 iterations are required to get a good approximation of the best value $x_{gc}$, and our binary search approach provably offers a linear convergence.}

\Comment{Is the method available?}
\Answer{TODO (Jerome)}

\subsection{Reviewer \#2}

\Comment{1) The $x^{\gc(s)}$ in equation (1) seems to me a novel contributing term that is of high importance. Perhaps its introduction should be highlighted more in the text.}
\Answer{We tried to make this point clearer in our revised version. However, it should be noted that this contribution is not a goal in itself, but rather a mean to perform stochastic sampling with respect to the weighted distribution. This distribution has all sorts of desirable mathematical properties, among which the fact that induced distributions for additive features (here the \GCContent) are nicely concentrated, allowing for a rejection-based approach.}

@@

\Comment{3-4) Typos\ldots}
\Answer{We fixed those typos.}

\subsection{Reviewer \#3}

\Comment{(1) There are many points where you iterate over "all sequences of length $|S|$" or all of those which can fold into $S$; it is not always clear. Why not introduce $F(s)$ for the folding space of a sequence, and $F^{-1}(S)$ for the inverse folding space of a structure, and simply use $F^{-1}(S)$ where this is what you mean?}
\Answer{While we agree that this proposition would lead to compact and well-defined mathematical statements, we prefer a implicit definition of the compatibility of a sequence to a given structure. Namely, we always allow every possible sequence of a length $|S|$ to be a member of the inverse folding space, possibly assigning some  energy penalty to it  (typically $+\infty$, or $\beta$) when some constraints are violated (restriction on allowed base-pairs, position-specific constraints...). The inverse folding space is therefore implicitly defined as the set of sequences having finite energy, and our algorithm is arguably more prone to generalization, e.g. to capture non-canonical base-pairs which are increasingly considered by the field of RNA bioinformatics.
}

\Comment{(2) You make the computation of $Z$ look like yet another set of DP recurrences, and as a consequence, you need a proof to show that it is linear in $n$. Surely, the reader wonders at that point how you can iterate over $(i,j)$ and compute an $(n,n)$ table in less than quadratic time. This mis-led understanding is created by the Nussinovic style of the recurrences.
 What you actually compute is a sparse, tree-shaped segment of that $(i,j)$-table.
 Why not represent the given structure as a tree of Empty, Singleton, Pair and Bifurcation nodes, such as 
\begin{align*}
  S<\texttt{.},B<P<\texttt{(},S<\texttt{.},E>,\texttt{)},P<\texttt{(},Em,\texttt{)}>>> \text{ for {\tt .(.)()}}  ? 
\end{align*}
 (You do not even need to explicitly represent the dots and brackets at the leaves.)\\
 Then, computation of $Z$ becomes just structural recursion on the tree,
     $Z(a,b, t) =$ defined by case analysis on t, such that, for example,
     $Z(a,b,P<'(',r,')'>)$ calls on $Z(a',b' r)$ for all $a'$ and $b'$.
And the result of the computation is just the tree annotated with the $Z_{ab}$ values. 
This way it is more transparent what is going on, and it is obvious that space and time are linear in $n$, because the size of the tree is in $O(n)$.}
\Answer{Great suggestion. However, adopting a pure tree-like representation would require elaborating on the correspondence between trees and structures, and might lead to technicalities in the context of stacking base-pairs. Therefore, we completely rewrote our equations, figure and algorithm to adopt a classic dot-parenthesis representation. While the end result may be less elegant than expected with a tree-like representations, this rewrite has the merit of dropping the double subscripts, hopefully conveying a better intuition of the principles underlying the complexity claim in our original manuscript.}

@@

\Comment{(3) Figure 1 is also a bit misleading. Make it clear that the stochastic choice is NOT made over the three cases depicted. The case is always fixed for given $i$, just the bases are chosen.}
\Answer{Absolutely. We changed this figure, which now explicitly distinguishes between the three different cases (different left-hand side diagrams), and also features the base emission probabilities.}

\Comment{(4) Section 2.2.3 is titled "Rejecting unsuitable candidates". It explains that candidates tend to be quite suitable, but somewhere, you should actually tell who is rejected and when ... }
\Answer{We completely rewrote this section, which now starts by explicitly reminding the problem addressed by the algorithm.}

\Comment{(5) recurring: "free-nucleotides in loops" -- isn't that the same, being "free" and in a loop? And: drop the hyphen.}
\Answer{The term of loop is sometimes used in a broader sense, which could be informally summarized as "a closed cyclic sequence of bases, connected either by the backbone or by base-pairs" (such that each connection but one goes in the $5'\to3'$ direction). This alternate definition was already used in the earlier works of M. Zuker, and maybe goes further back in the prehistory of RNA bioinformatics. Under this definition,  some of the bases involved in a loop may be paired, therefore we use the adjective to lift any ambiguity.
}

### 2013-03-09 ISMB/results_ISMB.tex
some stuff in src
method over (except typos)

\subsection{Design}
 To benchmark our method, we first started by sampling about $1000$ sequences per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

@@

Also, a critical propertie in RNA sequence design is 
the frequency of the MFE. 
The sequence should be in its target conformation long enough to
perform the desired action. Presented in Figure~\ref{fig:freq} we see that
there is a slow decline of the frequency with the increase in size. Yet,
for an average \GCContent, the frequency stays over $10\%$ even
at size of $100$ nucleotides.

>>>>

\subsection{Design}
 To benchmark our method, we first started by sampling about $100$ sequences per structure. After correcting the free nucleotides with
 \texttt{RNAinverse} we computed their MFE with \texttt{RNAfold} from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

@@

Also, a critical property in RNA sequence design is 
the frequency of the MFE. 
The sequence should be in its target conformation long enough to
perform the desired action. Presented in Figure~\ref{fig:freq} we see that
there is a slow decline of the frequency with the increase in size. Yet,
for an average \GCContent, the frequency stays over $10\%$ even
at size of $100$ nucleotides.
